---
title: "ggplot part 2"
author: "Amanda Mae Woodward"
date: "2024-10-22"
output: html_document
---

# Learning Outcomes
By the end of this class, students should be able to: 
- Explain how to alter information in the legend 
- Describe how to use facet wrap
- Explain how to export graphs


### Learning Outcome 1: Explain how to alter information in the legend
It can also be helpful to understand how to modify your legend: 
`+ theme(legend.position="none")` will remove the legend. You can also choose to move the legend to the top or bottom of the graph

Let's try this on a graph from the diamonds dataset: 
```{r}
library(ggplot2)
data(diamonds)
ggplot(diamonds, aes(color, price, fill= cut))+ 
  geom_boxplot()
```

We can also change the location: 
```{r}
ggplot(diamonds, aes(color, price, fill= cut))+ 
  geom_boxplot()+
  theme(legend.position = "bottom") ##positions are top, left, right, bottom, or none
```

To Change the legend title, you can use a couple of different codes, depending on the type of data: 
`scale_color_manual(name="Title")` works for scatter plots filled in with a factor
`scale_fill_manual(name="Title)` should work for a bar graph filled in by a factor
`scale_color_continuous()` and `scale_fill_continuous()` should work for a numeric fill
```{r}
ggplot(diamonds, aes(color, price, fill= cut))+ 
  geom_boxplot()+
  theme(legend.position = "bottom")+ 
  scale_fill_manual(name= "Cut of Diamond", values = c("#D7F2BA", "#0D2818", "#BC4749", "#058C42", "#FEC196"))
```

### Practice:

1) make a graph of bill length and bill depth. Include island as a third variable
```{r}
library(palmerpenguins)
data(penguins)
ggplot(penguins, aes(bill_length_mm, bill_depth_mm, color= island))+ 
  geom_point()
```

2) Move the legend to the top of the graph
```{r}
ggplot(penguins, aes(bill_length_mm, bill_depth_mm, color= island))+ 
  geom_point()+
  theme(legend.position = "top")
```

3) Remove the legend from the graph 
```{r}
ggplot(penguins, aes(bill_length_mm, bill_depth_mm, color= island))+ 
  geom_point()+
  theme(legend.position = "none")
```

4) Change the title of the legend
```{r}
ggplot(penguins, aes(bill_length_mm, bill_depth_mm, color= island))+ 
  geom_point()+
  theme(legend.position = "bottom")+
  scale_color_manual(name= "Island", values = c("#BC4749", "#058C42", "#FEC196"))
```

5) make a graph of bill length and bill depth. Include flipper length as a third variable
```{r}
ggplot(penguins, aes(bill_length_mm, bill_depth_mm, color = flipper_length_mm))+ 
  geom_point()
```

### Learning Outcome 2: Describe how to use facet wrap
Facet Wrap allows us to incorporate a third variable in another way. Specifically, it panels graphs so that we can make an individual graph for each level of a factor.

we need to make a graph and save it 
```{r}
ggplot(diamonds, aes(carat, price, color= cut))+ 
  geom_point()+
  geom_smooth(method= "lm", color= "black")+
  facet_wrap(~color) ##four variables

ggplot(diamonds, aes(color, price, colour = cut))+ 
  geom_boxplot()+
  facet_wrap(~cut) #three variables
```

### change labels on the axes
change levels beforehand
```{r}
levels(penguins$species)<- c("Adelie Penguin", "Chinstrap Penguin", "Gentoo Penguin")

ggplot(penguins, aes(bill_length_mm, bill_depth_mm))+ 
  geom_point()+ 
  facet_wrap(~species)
```

make a vector called labels. use `labeller=as_labeller())` in `facet_wrap()`
```{r}
speciesTitle<- c("Adelie Penguin"= "Adelie", "Chinstrap Penguin"= "Chinstrap", "Gentoo Penguin"= "Gentoo") #write current name in dataset = name you want to present

ggplot(penguins, aes(bill_length_mm, bill_depth_mm))+ 
  geom_point()+ 
  facet_wrap(~species, labeller= as_labeller(speciesTitle))
```

### Practice: 
1) create a facet wrap graph to generate a scatterplot of mpg and hp based on the number of cylinders a car has
```{r}
data(mtcars)
library(viridis)
cylinders<- c("4"= "Four Cylinder", "6"= "Six Cylinder", "8"= "Eight Cylinder")
ggplot(mtcars, aes(mpg, hp))+
  geom_point(color= "lightpink")+
  facet_wrap(~cyl, labeller= as_labeller(cylinders))+
  labs(title = "Miles per Gallon and Horsepower by Cylinder", x= "Miles per Gallon", y= "Horsepower")+
  theme_bw()
```

### Learning Outcome 3: Explain how to export graphs
to export graphs, I use the command `ggsave()`
Specifically, you can use the following arguments: 
`ggsave("name of graph", width, height, units, dpi)`
```{r}
getwd()
```
I don't usually specify a path in ggsave, but make sure my r code is in the directory I'd like it to be saved in. you can check your working directory by using this code
```{r}
ggsave("FinalMtcarsGraph.tiff", width = 10, height = 8, units = "in", dpi = 300)
```
you can change your working directory using this code: 
```{r}
setwd()
```
